---
title: "Image Data Summary"
output:
  html_document: default
  pdf_document: default
date: "2024-06-06"
---


```{r setup packages, include=FALSE}

library(tidyverse)
library(janitor)
library(renv)
library(knitr)
library(pals) # bunch of distinct color palettes
library(ggmap)
library(ggrepel)

# sequence of 25 colors that are mostly distinguishable, thanks
c25 <- c(
  "dodgerblue2", "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)
#pie(rep(1, 25), col = c25)

# polychrome in pals is also a good one, n =36

# turns out we have 37+ so
library(randomcoloR)
colors <- randomcoloR::distinctColorPalette(k = 40)


mytheme <- theme_classic()+ 
  ylab("Number of Sightings")+
  xlab("Sites North to South")+
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 7),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.title.x = element_text(size = 10))

```
# June 2024 Summary of Hollister Ranch Coastal Wildlife Camera Trapping Study

### Based on our conversation from last week, I've prepared the following summary figures plus some other options:
1) GIS map with all Hollister cameras and JLDP cameras (draft here, will make final in ArcGIS) 
2) pie chart for each camera separately on that so you can see variation across the coast
3) animal activity over month - stacked bar graph per month that has wild animals per camera night operating per month.   
4) relationship between people and diversity or abundance of wildlife
5) Other variables of interest?

```{r check out data, echo=FALSE, message=FALSE, warning=FALSE}

animl_wide <- read_csv("animldata_10jun2024.csv")
#str(animl)

animl <- animl_wide %>% 
  clean_names() %>% 
  pivot_longer(cols = 13:ncol(.), # columns containing species names, until the end of the df
               names_to = "species",
               values_to = "count"
               ) %>% 
  filter(count != 0) %>% # remove the species that have a 0 count because they weren't detected (keeps "empty" = 1)
  select(!c(id, camera_id, project_id, deployment_id, deployment_timezone)) %>%  # only keep useful columns
  
  mutate(deployment_name = str_remove(deployment_name, "-default")) %>% 
  mutate(deployment_name =  
         case_when(
           str_detect(deployment_name, "Alegria") ~ "Alegria",
           str_detect(deployment_name, "Caliente") ~ "Caliente",
           str_detect(deployment_name, "Gato") ~ "Gato",
           str_detect(deployment_name, "Llegua") ~ "Llegua",
           deployment_name == 864839041733151 ~ "Little Drake's",
           deployment_name == 864839042286761 ~ "Bulito",
           deployment_name == 864839040209047 ~ "Rennie's",
          deployment_name == 864839041745650 ~ "Auggie's",
           .default = as.character(deployment_name)
           )) %>% 
  mutate(date_time_original = ymd_hms(date_time_original)) %>% # giving a warning that 4 failed to parse ... checked and they're empties in teh middle of the april at auggie's so deleting below
  filter(!is.na(date_time_original)) %>% 
  separate(date_time_original, into = c('date', 'time'), sep=' ', remove = TRUE) %>% # make columns for date and time separated
  mutate(year = year(date)) %>% # makes this recognized as year format
  mutate(month = month(date, label = TRUE))  #same
  

#105,000 + entries, whew
```

## Draft Map of all Hollister Ranch Cameras
### (final draft will include Dangermond cameras)

```{r map, echo=FALSE, message=FALSE, warning=FALSE}

###################### pull in hr camera data ###########################
cam_sites <- read_csv("hroa_cam_sites.csv") %>% 
  mutate(WKT = str_remove(WKT, "POINT \\(")) %>% 
  mutate(WKT = str_remove(WKT, "\\)")) %>% 
  separate(WKT, c("lon", "lat"), sep = " ") %>% #separate weird coords into lon lat
  mutate(lon = as.numeric(lon)) %>% 
  mutate(lat = as.numeric(lat)) %>% 
  select(lat, lon, name) %>% 
  filter(name != "Surfin Cowboys") %>% 
  filter(name != "Bulito Beach Access") %>% 
  filter(name != "John's Pond Culvert") %>% 
  filter(name != "Drake's Access Point") %>% 
  mutate(name = case_when(
    name == "Rennie's / North Bulito Pond" ~ "Rennie's",
    name == "Sacate/Coyote" ~ "Sacate",
    name == "Bulito South" ~ "Bulito",
    .default = as.character(name)
  ))

############# Set map parameters and make the base map ################
register_google(key = "AIzaSyDMo_xQFMKeU320T6-fnJgvQ4KwzoLubgA") # new key june 2024

mapbase <- get_map(location = c(lon = -120.35, lat = 34.5), 
                   zoom = 12,
                   maptype = "satellite", 
                   source = "google")

#basemap <- ggmap(mapbase) 

#basemap # to make sure it's in the right place
#######################################################################

#lon_crop = c(120.5, 120.2)
#lat_crop = c(34.8, 34.2)

cammap <- ggmap(mapbase) +
  geom_point(data = cam_sites, 
             aes(x = lon, y = lat),
             colour = "lightblue")+
  geom_label_repel(data = cam_sites,  # need to use ggrepel because the stupid labels overlap
             aes(x = lon, y = lat, label = name), 
             #nudge_x = 0.01,
             #nudge_y = 0.01,
             size = 2,
             force = 2, # makes labels less likely to overlap
             min.segment.length = unit(0, "lines") #adds a line to point
             )+
  # want a specific part of the map only 
  scale_y_continuous(limits = c(34.420, 34.50))+

  theme_bw()

cammap

```


## Summarizing the data for some quick visualizations
```{r making summary tables, echo=FALSE, message=FALSE, warning=FALSE}

########################################################################################

######### make table of total images by deployment_name ###########
count_site <- animl %>%
  count(deployment_name) %>% 
  rename(n_images_total = n) # rename n to be more informative


########################################################################################

######### make table of animal activity ( = incidences) by species summed across all deployment_ids ###########
count_species <- animl %>%
  count(species) %>%  # makes a summary table of # images per species per deployment_id
  rename(n_images = n) # rename n to be more informative


########################################################################################

######### make table of animal activity ( = incidences) by species and deployment_ids ###########
count_species_site <- animl %>%
  count(deployment_name, species) %>%  # makes a summary table of # images per species per deployment_id
  rename(n_images = n) # rename n to be more informative


########################################################################################

######### make table of TOTAL wild animal activity ( = incidences that weren't empty, people, or domestic animals) summed across deployment_ids ###########
count_activity <- animl %>%
  filter(species != "empty") %>% 
  filter(species != "person")%>% 
  filter(species != "human")%>% 
  filter(species != "dog")%>% 
  filter(species != "domestic dog")%>% 
  filter(species != "equestrian") %>% 
  count(deployment_name) %>%  # makes a summary table of # images per species per deployment_id
  rename(n_animal_imgs = n) # rename n to be more informative


########################################################################################

######### make table of total human activity ( = incidences that weren't empty, people, or domestic animals) summed across deployment_ids ###########
human_activity <- animl %>%
  filter(species == "person"|species == "human"|species == "dog"|species == "domestic dog"|species == "equestrian") %>% 
  count(deployment_name) %>%  # makes a summary table of # images of humans per deployment_id
  rename(n_human_imgs = n) # rename n to be more informative

########################################################################################

######### make table number of wildlife species at each site ###########
num_species_site <- animl %>%
  filter(species != "empty") %>% 
  filter(species != "person")%>% 
  filter(species != "human")%>% 
  filter(species != "dog")%>% 
  filter(species != "domestic dog")%>% 
  filter(species != "equestrian") %>%  
  group_by(deployment_name) %>%  # makes a summary table of # species per deployment_id
  summarize(num_species = n_distinct(deployment_name, species)) # how many unique species at each site
  
########################################################################################


######### make table one big summary table by site ###########

camera_activity_summary <- full_join(count_site, count_activity,  by = join_by(deployment_name)) %>% 
  full_join(human_activity, by = join_by(deployment_name)) %>%
  full_join(num_species_site, by = join_by(deployment_name)) %>% 
  mutate(percent_animals = (n_animal_imgs/n_images_total)*100) %>% 
  mutate(deployment_name = fct_relevel(deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente"))) %>%  # relevel sites so they're in a N > S order
  arrange(deployment_name)

kable(camera_activity_summary
      ,col.names = c("Camera Location (N to S)", "Total Images Taken" , "Number of Animal Sightings", "Number of Human Sightings", "Number of Unique Species", "Percent of Images Containing Animals")
      )
  
```


```{r visualizations of data over time, echo=FALSE, message=FALSE, warning=FALSE}

## Temporal visualizations are more tricky because some of the cameras have date errors
### for now we will just filter incorrect dates out

animl_time <- animl %>% 
  filter(year > 2018)

######### make table of trap nights by deployment_name ###########

# trap night is defined as every night a camera is active, so this isn't really trap nights, its "this camera was triggered nights"

trigger_nights_site <-  animl_time%>% 
  group_by(deployment_name) %>% 
  summarize(trigger_nights = n_distinct(deployment_name, date)) # how many unique dates at each site

trapnights_site <- animl_time %>% 
  group_by(deployment_name) %>% 
  summarize(start = min(date),
            end = max(date),
            trapnights_w_photos = n_distinct(date),
            trapnights = difftime(max(date), min(date), units = "days")
            )
#view(trapnights_site)
########################################################################################


######### make table of species by month ###########
species_month <- animl_time %>%
  count(month, year, species) %>%  # makes a summary table of # images per species per month
# includes count of empties
  rename(n_imgs = n) # rename n to be more informative

#species_site
########################################################################################

######### make table of animal activity (seconds = incidences) by deployment_id by date ###########
count_site_date <- animl %>%
  count(deployment_name, month) # makes a summary table of # animal sightings per species per deployment_id per month
########################################################################################


######### make table of total images summed across all deployment_ids by date ###########
count_month <- animl %>%
  count(month) # makes a summary table of # animal-seconds per month
########################################################################################


######### make table of animal activity ( = incidences) summed across all deployment_ids by date ###########
count_month <- animl %>%
  count(month) # makes a summary table of # animal-seconds per month
########################################################################################

```


## Some plots to summarize preliminary data

### Total nights that cameras were active at each location
```{r trap nights by site, echo=FALSE, message=FALSE, warning=FALSE}

trapnights_site$deployment_name <- fct_relevel(count_site$deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente")) # relevel sites so they're in a N > S order

nightsbysite <- ggplot(trapnights_site, aes(deployment_name, trapnights)) +
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  mytheme +
  ylab("Number of Nights Camera was Active")+
  xlab("Camera Location - North to South")

nightsbysite

```

### Total images taken at each location (including animal images, human images, and blanks)

```{r count by site, echo=FALSE, message=FALSE, warning=FALSE}
count_site$deployment_name <- fct_relevel(count_site$deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente")) # relevel sites so they're in a N > S order

countbysite <- ggplot(count_site, aes(deployment_name, n_images_total)) +
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  mytheme +
  ylab("Number of Total Images Taken")+
  xlab("Camera Location - North to South")

countbysite
```

### Percent of images that contained animals at each location (this corrects for different numbers of images taken at different locations, better for comparing sites to each other)
```{r percent animals by site, echo=FALSE, message=FALSE, warning=FALSE}

camera_activity_summary$deployment_name <- fct_relevel(count_site$deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente")) # relevel sites so they're in a N > S order

percentbysite <- ggplot(camera_activity_summary, aes(deployment_name, percent_animals)) +
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  mytheme +
  ylab("Percent of Images Taken Containing Wildlife")+
  xlab("Camera Location - North to South")

percentbysite
```

### What wildlife species were seen across the whole ranch and how many times?
```{r frequency of species, echo=FALSE, message=FALSE, warning=FALSE}

count_species_plot <- count_species %>% # make a new version of this table with wildlife sp only
  filter(species != "empty") %>% 
  filter(species != "animal") %>% 
  filter(species != "person") %>% 
  filter(species != "domestic_dog") %>% 
  filter(species != "equestrian") %>% 
  filter(species != "human") %>% 
  filter(species != "dog") %>% 
  filter(!str_detect(species, "unidentified"))%>% 
  filter(!str_detect(species, "dead"))

countbyspecies <- ggplot(count_species_plot, aes(x = reorder(species, -n_images)
                                                 , y = n_images)) + # this reorders them from biggest to smallest i think
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  mytheme +
  ylab("Number of Sightings")+
  xlab("Species")

countbyspecies
```

### How many times was each species seen at each location?
```{r stacked bar chart of species by site, echo=FALSE, message=FALSE, warning=FALSE}

species_site_plot <- count_species_site %>% # make a new version of this table with wildlife sp only
  filter(species != "empty") %>% 
  filter(species != "animal") %>% 
  filter(species != "person") %>% 
  filter(species != "domestic_dog") %>% 
  filter(species != "equestrian") %>% 
  filter(species != "human") %>% 
  filter(species != "dog") %>% 
  filter(!str_detect(species, "unidentified")) %>% 
  filter(!str_detect(species, "dead"))

# set order of sites N to S
species_site_plot$deployment_name <- deployment_name <- fct_relevel(species_site_plot$deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente")) # relevel sites so they're in a N > S order

stacked_spp_site <- ggplot(species_site_plot, aes(deployment_name, n_images)) +
  geom_bar(aes(fill = species), # lets me make clustered bar chart
           stat = "identity", # idk but necessary
           #position = "dodge" # otherwise default is stacked barchart
           )+
  scale_fill_manual(values = unname(colors))+ # not sure why colors need to be unnamed
  mytheme +  
  ylab("Number of Sightings")+
  xlab("Camera Locations - North to South")
  

stacked_spp_site

```

### How *frequently* was each species seen at each location, relatively?
#### These pie charts use relative frequency (percent of total wildlife sightings) to correct for differences in detection between cameras
 
```{r wildlife frequency pie charts, echo=FALSE, message=FALSE, warning=FALSE}

# don't really need ot change much, just swap stacked barchat to position = fill to make it percents instead of counts, use coord_polar() to make it a donut, and facet wrap by deployment_name

# set order of sites N to S
species_site_plot$deployment_name <- fct_relevel(species_site_plot$deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente")) # relevel sites so they're in a N > S order

# plot
species_site_freq <- ggplot(species_site_plot, aes(x = 4, y = n_images, fill = species)) +
  geom_bar(position = "fill", stat = "identity")+
  facet_wrap(~deployment_name)+
  coord_polar(theta = "y")+
  scale_fill_manual(values = unname(colors))+ # not sure why colors need to be unnamed
  theme_void()

species_site_freq

```


 
### How many times were humans and domestic animals seen at each location?
```{r humans and dogs, echo=FALSE, message=FALSE, warning=FALSE}

human_site <- count_species_site %>% # make a new version of this table with human + domestic sp only
  filter( species =="person"| species == "domestic_dog"| species ==  "equestrian"| species == "human"| species == "dog") %>% 
  mutate(species = case_when(species == "human" ~ "person",
                             .default = as.character(species)))

# set order of sites N to S
human_site$deployment_name <- deployment_name <- fct_relevel(human_site$deployment_name, c("Gato", "Kayak (St. Augustine West)", "Llegua" , "Auggie's", "Rennie's", "Bulito", "Panochas", "Little Drake's", "Sacate", "Cuarta", "Alegria", "Caliente")) # relevel sites so they're in a N > S order

human_site_plot <- ggplot(human_site, aes(deployment_name, n_images)) +
  geom_bar(aes(fill = species), # lets me make clustered bar chart
           stat = "identity", # idk but necessary
           #position = "dodge" # otherwise default is stacked barchart
           )+
  mytheme +
  ylab("Number of Sightings")+
  xlab("Camera Location - North to South")


human_site_plot

```
 
### Wildlife sightings per month, broken down by species
```{r monthly wildlife, echo=FALSE, message=FALSE, warning=FALSE}

wildlife_month <- species_month %>% 
  filter(species != "empty") %>% 
  filter(species != "animal") %>% 
  filter(species != "person") %>% 
  filter(species != "domestic_dog") %>% 
  filter(species != "equestrian") %>% 
  filter(species != "human") %>% 
  filter(species != "dog") %>% 
  filter(!str_detect(species, "unidentified")) %>% 
  filter(!str_detect(species, "dead")) %>% 
  unite("month_year", month, year, sep = " ")

wildlife_month$month_year <- fct_relevel(wildlife_month$month_year, c( "Jun 2023","Jul 2023", "Aug 2023", "Sep 2023", "Oct 2023", "Nov 2023", "Dec 2023", "Jan 2024", "Feb 2024", "Mar 2024", "Apr 2024", "May 2024", "Jun 2024"   ))

 monthly_spp <- ggplot(wildlife_month, aes(month_year, n_imgs)) +
  geom_bar(aes(fill = species), # lets me make clustered bar chart
           stat = "identity"
           )+
  scale_fill_manual(values = unname(colors))+ # not sure why colors need to be unnamed
  mytheme +  
  ylab("Number of Sightings")+
  xlab("Month and Year")

monthly_spp

```


### Is there a relationship between people sightings and wildlife sightings?

```{r wildlife count ~ human count, echo=FALSE, message=FALSE, warning=FALSE}

humanxwildlife_scatter <- ggplot(camera_activity_summary, aes(x = n_human_imgs, y = n_animal_imgs))+
  geom_point() +
  mytheme +
  ylab("Number of Animal Sightings")+
  xlab("Number of Human Sightings")

humanxwildlife_scatter

```


### Is there a relationship between people sightings and wildlife species richness (# species)?

```{r richness ~ human count, echo=FALSE, message=FALSE, warning=FALSE}

humanxrichness_scatter <- ggplot(camera_activity_summary, aes(x = n_human_imgs, y = num_species))+
  geom_point() +
  mytheme +
  ylab("Number of Wildlife Species")+
  xlab("Number of Human Sightings")

humanxrichness_scatter

```

#### Seems like a slight positive correlation between wildlife sightings and human sightings, which is likely a result of camera performance, e.g. some cameras take more pictures and just capture more humans and more wildlife.